---
title: "5361_stat_hw_Qi_You"
output:
  pdf_document: default
  html_document: default
---

## Question 1b
```{r setup, include=FALSE}
target <- function(x,sita){
  n=length(x)
  value <--n*log(pi)
  for (i in 1:n){
    value <- value-log(1+(sita-x[i])^2)
  }
  return(value)
}
x <- c(1.77, -0.23, 2.76, 3.80, 3.47, 56.75, -1.34, 4.24, -2.44,
       3.29, 3.71, -2.40, 4.53, -0.07, -1.05, -13.87, -2.53, -1.75)


derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

derivitive2 <- function(s,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*((1-(sita-x[i])^2)/(1+(sita-x[i])^2)^2)
  }
  return(value)
}


estimate <- function(initilal_value){
  sita <- array()
  sita[1] <- initilal_value
  i <- 1
  difference <- 100
  while(abs(difference)>10^(-4)&i<100){
    sita[i+1] <- sita[i]-derivitive1(x,sita[i])/derivitive2(x,sita[i])
    difference <- target(x,sita[i+1])-target(x,sita[i])
    i <- i+1
  }
  plot(sita,type='l',main=initilal_value,xlab = 'times')
}
tempt <- seq(-2,4,by=0.1)
plot(tempt,target(x,tempt),type = 'l')
start_points <- c(-11, -1, 0, 1.5, 4, 4.7, 7, 8)

for(i in 1:length(start_points)){
  estimate(start_points[i])
}

estimate(mean(start_points))

```
## Equations
\begin{align}
l(\theta)\\
          &= \ln\prod_{i = 1}^n \frac{1}{\pi[1+(x-\theta)^2]}&\\
          &= \sum_{i=1}^n \ln\frac{1}{\pi[1+(x-\theta)^2]}&\\
          &= \sum_{i=1}^n [\ln\frac{1}{\pi} + \ln\frac{1}{1+(x-\theta)^2}]&\\
          &= -n\ln\pi - \sum_{i=1}^n \ln[1+(x-\theta)^2]&\\
\end{align}

\begin{align}
l' (\theta)\\
          &= 0 - \sum_{i=1}^n \frac{2(\theta-x_i)}{1+(\theta-x_i)^2}&\\
          &= - 2\sum_{i=1}^n \frac{\theta - x_i}{1+(\theta-x_i)}&\\
\end{align}

\begin{align}
l''(\theta)\\
          &= -2\sum_{i=1}^n \frac{1+(\theta-x_i)^2-2(\theta-x_i)(\theta-x_i)}{[1+(\theta-x_i)^2]^2}&\\
          &= -2\sum_{i=1}^n \frac{1-(\theta-x_i)^2}{[1+(\theta-x_i)^2]^2}&\\
\end{align}

We can set x to be tant(x). In this way we can compute the value of I easiliy
\begin{align}
I(\theta)\\
          &= n\int\frac{\{p'(x)\}^2}{p(x)}dx\\ 
          &= n\int\frac{4(x-\theta)^2}{\pi[1+(x-\theta)^2]^4}*\pi[1+(x-\theta)^2]dx&\\
          &= \frac{4n}{\pi} \int_{-\infty}^\infty \frac{(x-\theta)^2}{[1+(x-\theta)^2]^3}dx&\\
          &= \frac{4n}{\pi} \int_{-\infty}^\infty \frac{x^2}{(1+x^2)^3}dx&\\
          &= \frac{4n}{\pi} \int_{-\infty}^\infty [(\frac{1}{(1+x^2)^2}-\frac{1}{(1+x^2)^3})]dx&\\
          &= \frac{4n}{\pi} (\int_{-\infty}^\infty \frac{1}{(1+x^2)^2}dx-\int_{-\infty}^\infty\frac{1}{(1+x^2)^3}dx)&\\
          &= \frac{4n}{\pi} [\int_{-\infty}^{\infty}\frac{1}{(1+x^2)^2}dx-(\frac{x}{4(x^2+1)^2}|_{-\infty}^{\infty}+\frac{3}{4}\int_{-\infty}^{\infty}\frac{dx}{(x^2+1)^2})]&\\
          &= \frac{4n}{\pi}(\int_{-\infty}^{\infty}\frac{1}{4(1+x^2)^2}dx-\frac{x}{4(x^2+1)^2}|_{-\infty}^{\infty})&\\
          &= \frac{4n}{\pi} [\frac{1}{4}(\frac{x}{2(x^2+1)}|_{-\infty}^{\infty}+\frac{1}{2}\int_{-\infty}^{\infty}\frac{1}{1+x^2}dx)-\frac{x}{4(x^2+1)^2}|_{-\infty}^\infty]&\\
          &= \frac{4n}{\pi}(\frac{x(x^2-1)}{8(x^2+1)^2}|_{-\infty}^\infty+\frac{1}{8}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}} \frac{\sec^2t}{1+\tan^2t}dt)&\\
          &= \frac{4n}{\pi}(0+\frac{\pi}{8})&\\
          &= \frac{n}{2}&\\
\end{align}

## Functions I setup

```{r derivitive1&2}
derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

derivitive2 <- function(s,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*((1-(sita-x[i])^2)/(1+(sita-x[i])^2)^2)
  }
  return(value)
}
```

## Plots(1b)

Here are the plots of my result 

```{r ddd, echo=FALSE}
plot(tempt,target(x,tempt),type = 'l')
```

Here are the plots of mutiple start points
```{r multiple}
for(i in 1:length(start_points)){
  estimate(start_points[i])
}
```

This is the plot of mean
```{r mean}
estimate(mean(start_points))
```

## Question 1(c)
This time we use fixed points
```{r kkk, include=FALSE}
x <- c(1.77, -0.23, 2.76, 3.80, 3.47, 56.75, -1.34, 4.24, -2.44,
       3.29, 3.71, -2.40, 4.53, -0.07, -1.05, -13.87, -2.53, -1.75)


derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

G <- function(alpha,x,sita){
  value <- alpha*derivitive1(x,sita)+sita
  return(value)
}


estimate <- function(start_points,alpha){
  sita <- array()
  sita[1] <- start_points
  difference <- 100
  i <- 1
  while(abs(difference)>10^(-4)&i<100){
    sita[i+1] <- G(alpha,x,sita[i])
    difference <- sita[i+1]-sita[i]
    i <- i+1
  }
  plot(sita,type='l',main=c(start_points,alpha),xlab='times')
}
start_points <- c(-11, -1, 0, 1.5, 4, 4.7, 7, 8)
alpha <- c(1, 0.64, 0.25)
for(i in 1:length(alpha)){
  for(j in 1:length(start_points)){
    estimate(start_points[j],alpha[i])
  }
}
```

These are functions I established to compute derivition
```{r cars}
derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

G <- function(alpha,x,sita){
  value <- alpha*derivitive1(x,sita)+sita
  return(value)
}

```

## Here are plots1(c)
These plots are using different alphas and start points

```{r pressure, echo=FALSE}
for(i in 1:length(alpha)){
  for(j in 1:length(start_points)){
    estimate(start_points[j],alpha[i])
  }
}
```

## Question 1(d)
This section we use Fisher Scoring to find MLE, with I(theta) being constant of n/2
```{r setup1, include=FALSE}
x <- c(1.77, -0.23, 2.76, 3.80, 3.47, 56.75, -1.34, 4.24, -2.44,
       3.29, 3.71, -2.40, 4.53, -0.07, -1.05, -13.87, -2.53, -1.75)

derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

I <- length(x)/2

estimate <- function(start_points){
  sita <- array()
  sita[1] <- start_points
  difference <- 100
  i <- 1
  while (abs(difference>10^(-4)&i<100)) {
    sita[i+1] <- sita[i]+derivitive1(x,sita[i])/I
    difference <- sita[i+1]-sita[i]
    i <- i+1
  }
  plot(sita,type='l',main=start_points,xlab='times')
}
start_points <- c(-11, -1, 0, 1.5, 4, 4.7, 7, 8)
for(i in 1:length(start_points)){
  estimate(start_points[i])
}
```


The functions we set up are similar with previous functions, and we compute I
```{r cars1}
derivitive1 <- function(x,sita){
  value <- 0
  for (i in 1:length(x)){
    value <- value-2*(sita-x[i])/(1+(sita-x[i])^2)
  }
  return(value)
}

I <- length(x)/2
```

## Plots1(d)

These are plots that I use multiple start points

```{r pressure1, echo=FALSE}
for(i in 1:length(start_points)){
  estimate(start_points[i])
}
```

## 1(e)

With observation of the results and plots, we can see that Newton method has the highest speed and stability. For Fisher scoring, it has lower speed, but we can see from the plots, with alpha being some point the line is fluctuating up and down, so it has lowest stability. And for fixed point method, the speed of it is lowest but is relatively stable. But fixed-point has its disadvantage, not being able to utilize some points.

## Qusetion 2
```{r setup2, include=FALSE}
x <- c(3.91, 4.85, 2.28, 4.06, 3.70, 4.04, 5.46, 3.53, 2.28, 1.96,
     2.53, 3.88, 2.22, 3.47, 4.82, 2.46, 2.99, 2.54, 0.52)

theta <- seq(-pi,pi,by=0.001)

loglikelihood <- function(x,theta){
  value <- 0
  for (i in 1:length(x)){
    value <- value+log((1-cos(x[i]-theta))/(2*pi))
  }
  return(value)
}

value <- array()
for (i in 1:length(theta)){
  value[i] <- loglikelihood(x,theta[i])
}

plot(theta,value,type = 'l')
mm <- mean(x)
target <- asin(mm-pi)

derivetive1 <- function(x,theta){
  value <- 0
  for (i in 1:length(x)){
    value <- value+sin(x[i]-theta)/(1-cos(x[i]-theta))
  }
  return(value)
}

derivetive2 <- function(x,theta){
  value <- 0
  for (i in 1:length(x)){
    value <- value-1/(cos(x[i]-theta)-1)
  }
  return(value)
}


MLE <- function(theta){
  diff <- 1
  i <- 1
  while(abs(diff)>10^(-4)){
    theta[i+1] <- theta[i]-derivetive1(x,theta[i])/derivetive2(x,theta[i])
    diff <- theta[i+1]-theta[i]
    i <- i+1
  }
  return(theta)
}

theta <- array()
theta[1] <- target
s <- MLE(theta)
theta <- array()
theta[1] <- -2.7
ss <- MLE(theta)
theta <- array()
theta[1] <- 2.7
sss <- MLE(theta)

textt <- seq(-pi,pi,length.out=200)
m <- list()
for (i in 1:length(textt)){
  theta <- array()
  theta[1] <- textt[i]
  k <- MLE(theta)
  k <- data.frame(k)
  m[i] <- k
}

result <- array()
for (i in 1:length(m)){
  l <- m[[i]]
  result[i] <- l[length(l)]
}

j <- 1
num <- list()
for (i in 1:length(result)){
  if (abs(result[i]-result[j])>10^(-4)){
    j <- i
    num <- append(num,j-1)
   }
}
```

## The likelihood function
\begin{equation}
E[X|\theta] = \pi + \sin(\theta)
\end{equation}


## two points
```{r points}
target <- asin(mm-pi)
print(target)
target1 <- pi-target
print(target1)
```

## The sequence of same converge start points
```{r c2}
loglikelihood <- function(x,theta){
  value <- 0
  for (i in 1:length(x)){
    value <- value+log((1-cos(x[i]-theta))/(2*pi))
  }
  return(value)
}
for (i in 1:length(result)){
  if (abs(result[i]-result[j])>10^(-4)){
    j <- i
    num <- append(num,j-1)
   }
}
num <- unique(num)
print(unique(num[1:length(num)-1]))
```

## Plots2
We plot the value of our target

```{r p2, echo=FALSE}
plot(value,type = 'l',xlab='theta')
```

we also plot the picture with strat point being 'target'
```{r theta}
MLE <- function(theta){
  diff <- 1
  i <- 1
  while(abs(diff)>10^(-4)){
    theta[i+1] <- theta[i]-derivetive1(x,theta[i])/derivetive2(x,theta[i])
    diff <- theta[i+1]-theta[i]
    i <- i+1
  }
  plot(theta,type='l')
  return(theta)
}

theta <- array()
theta[1] <- target
s <- MLE(theta)
```

## Qustion 3(b)
```{r setup3, include=FALSE}


beetles <- data.frame(
  days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
  beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))

r <- 0.1
k <- 100
N0 <- 2
A_matrix <- function(r,k,times){
  A <- matrix(nrow=times,ncol=2)
  for (i in 1:times){
    A[i,1] <- (N0^2-N0^2*exp(-1*r*beetles$days[i]))/((N0+(k-N0)*exp(-1*r*beetles$days[i]))^2)
    A[i,2] <- beetles$days[i]*(k-N0)*exp(-1*r*beetles$days[i])*k*N0/((N0+(k-N0)*exp(-1*r*beetles$days[i]))^2)
  }
  A
}


Z_matrix <- function(r,k,times){
  z <- matrix(nrow=times,ncol=1)
  for (i in 1:times){
    z[i,1] <- beetles$beetles[i]-k*2/(2+(k-2)*exp(-r*beetles$days[i]))
  }
  z
}

theta <- list()
sss <- t(cbind(k,r))
theta[1] <- data.frame(sss)
temt <- theta[[1]]


error <- function(theta,i){
  n1 <- theta[[i+1]]
  n2 <- theta[[i]]
  error <- abs(n1[1]-n2[1])+abs(n1[2]-n2[2])
}
i <- 1
error1 <- 100
while(error1>10^(-4)){
  temt <- theta[[i]]
  s <- A_matrix(temt[2],temt[1],10)
  m <- Z_matrix(temt[2],temt[1],10)
  argue <- solve(t(s)%*%s)%*%t(s)%*%m
  theta[[i+1]] <- theta[[i]]+argue
  error1 <- error(theta,i)
  i <- i+1
}

NT <- function(r,k,i){
  value <- k*2/(2+(k-2)*exp(-r*beetles$days[i]))
}

errors <- function(r,k,times){
  zz <- Z_matrix(r,k,times)
  value <- 0
  for (i in 1:length(zz)){
    value <- zz[i]^2+value
  }
  return(value)
}

kk <- array()
for(i in 1:length(theta)){
  parameter <- theta[[i]]
  kk [i]<- errors(parameter[2],parameter[1],10)
}
plot(kk,type='l')
```

## Functions

Here are matrix we need to establish to build our function
```{r c3}
A_matrix <- function(r,k,times){
  A <- matrix(nrow=times,ncol=2)
  for (i in 1:times){
    A[i,1] <- (N0^2-N0^2*exp(-1*r*beetles$days[i]))/((N0+(k-N0)*exp(-1*r*beetles$days[i]))^2)
    A[i,2] <- beetles$days[i]*(k-N0)*exp(-1*r*beetles$days[i])*k*N0/((N0+(k-N0)*exp(-1*r*beetles$days[i]))^2)
  }
  A
}


Z_matrix <- function(r,k,times){
  z <- matrix(nrow=times,ncol=1)
  for (i in 1:times){
    z[i,1] <- beetles$beetles[i]-k*2/(2+(k-2)*exp(-r*beetles$days[i]))
  }
  z
}
```

## Sum of squared errors


```{r p3, echo=FALSE}
plot(kk,type='l',xlab='times')
```

## Question 3(c)
This time we need to optimize the value of r, K, sigma
```{r setup4, include=FALSE}


beetles <- data.frame(
  days    = c(0,  8,  28,  41,  63,  69,   97, 117,  135,  154),
  beetles = c(2, 47, 192, 256, 768, 896, 1120, 896, 1184, 1024))

times <- beetles$days
n <- beetles$beetles
u <- expression(-log((2*pi*sigma)^0.5)-((log(N)-log(k*2/(2+(k-2)*exp(-r*t))))^2)/(2*(sigma^2)))

d_sigma <- as.expression(D(u,'sigma'))
dd_sigma <- as.expression(D(d_sigma,'sigma'))

d_k <-as.expression(D(u,'k'))
dd_k <- as.expression(D(d_k,'k'))

d_r <- as.expression(D(u,'r'))
dd_r <- as.expression(D(d_r,'r'))

theta <- list()
theta[1]<- data.frame(c(1,1000,0.1))

m <- theta[[1]]
value <- 0
value1 <- 0
value2 <- 0
for(i in 1:length(times)){
  t=times[i]
  N=n[i]
  sigma=1
  k=1000
  r=0.1
  value <- value+eval(dd_sigma)
  value1 <- value1+eval(dd_k)
  value2 <- value2+eval(dd_r)

}

#m <- theta[[1]]
times <- beetles$days

get_d_sigma <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(d_sigma,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
  }
  return(kk)
}

get_dd_sigma <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(dd_sigma,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
    #kk[2] <- kk[2]+eval(dd_sigma,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
    #result <- kk[1]/kk[2]
  }
  return(kk)
}

get_d_k <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(d_k,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
  }
  return(kk)
}

get_dd_k <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(dd_k,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
    #kk[2] <- kk[2]+eval(dd_sigma,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
    #result <- kk[1]/kk[2]
  }
  return(kk)
}

get_d_r <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(d_r,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
  }
  return(kk)
}

get_dd_r <- function(m){
  kk <- 0
  for(i in 1:length(beetles$days)){
    kk <- kk+eval(dd_r,envir=list(sigma=m[1],k=m[2],r=m[3],t=times[i],N=n[i]))
  }
  return(kk)
}

result <- get_dd_k(m)
error <- function(theta,i){
  n1 <- theta[[i+1]]
  n2 <- theta[[i]]
  error <- sum(abs(n1 - n2))
  return(error)
}

difference <- 100
i <- 1
kF <- m[2]
rF <- m[3]
sigmaF <- m[1]
while(difference>0.000001&i<1000){
  m <- theta[[i]]
  kF <- kF-(get_d_k(m)/get_dd_k(m))
  rF <- rF-(get_d_r(m)/get_dd_r(m))
  sigmaF <- sigmaF-(get_d_sigma(m)/get_dd_sigma(m))
  theta[i+1] <- data.frame(c(sigmaF,kF,rF))
  updatek <- abs((get_d_k(m)/get_dd_k(m)))
  updater <- abs((get_d_r(m)/get_dd_r(m)))
  updatesigma <- abs((get_d_sigma(m)/get_dd_sigma(m)))
  #difference <- error(theta,i)
  difference <- updatek + updater + updatesigma
  i <- i+1
  
}
print(i)
print(kF)
print(rF)
print(sigmaF)

```

## Partial Derivative
```{r partial}
d_sigma <- as.expression(D(u,'sigma'))
dd_sigma <- as.expression(D(d_sigma,'sigma'))

d_k <-as.expression(D(u,'k'))
dd_k <- as.expression(D(d_k,'k'))

d_r <- as.expression(D(u,'r'))
dd_r <- as.expression(D(d_r,'r'))
```

## 3c

I use Newton method to optimize the r,K, and square of sigma.
These numbers are

Times we use 
```{r p4, echo=FALSE}
print(i)
```
The results
```{r results}
print(kF)
print(rF)
print(sigmaF)
```


